R code for 'Apparent physiological costs of blood parasite infection in the early-life of a vertebrate host - only in highest infections

### 2022-04-04 MS_physiology_infection.Rmd
Add files via upload

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5),size=5) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) + 
  scale_x_discrete(labels = c("First\nsampling","Second\nsampling","First\nsampling","Second\nsampling"))+  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 22, face = "bold"),
        legend.text = element_text(size = 20),
        legend.title = element_blank(),
        legend.position = c(0.85, 0.90),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text.x = element_text(size = 22, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 22, margin =  unit(c(0, 1, 0, 0), "mm")))

@@

## Figure 2b
## -----------------------------------------------------------------------------
Fig_2B <- ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1)) +
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey")) + 
  scale_x_discrete(labels = c("First\nsampling","Second\nsampling"))+
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 22),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 5, 0, 0), "mm")),
        axis.text.x = element_text(size = 22, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 22, margin =  unit(c(0, 1, 0, 3), "mm")),
        plot.margin = unit(c(0,0,0,30), "pt"))

@@

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.3),
            label = "P = 0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.1)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.1,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

@@

Fig_4C = ggplot(mod_bc_para) + 
  aes(as.factor(mod_bc_para$x),mod_bc_para$predicted) + 
  geom_hline(yintercept = mean(mod_bc_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc_para$lower_CI, ymax = mod_bc_para$higher_CI),
                colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc_para$predicted - mod_bc_para$se,
                    ymax = mod_bc_para$predicted + mod_bc_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),  
        axis.text = element_text(size = 16), 
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 60, xend = 5, yend = 60)) + 
    geom_text(aes(x = 3, y = 67),
              label = "P = 0.080", size = 5) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 60, xend = 1, yend = 53)) + 
  geom_segment(size = .8,
                 aes(x = 5, y = 60, xend = 5, yend = 53))
```

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4D = ggplot(mod_gr_para) + 
  aes(as.factor(mod_gr_para$x),mod_gr_para$predicted) + 
  geom_hline(yintercept = mean(mod_gr_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr_para$lower_CI,
                    ymax = mod_gr_para$higher_CI), colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr_para$predicted - mod_gr_para$se,
                    ymax = mod_gr_para$predicted + mod_gr_para$se),size = 8, 
                colour = wes_palette("Zissou1",5), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
               aes(x = 1, y = 30, xend = 5, yend = 30)) + 
  geom_text(aes(x = 3, y = 31),
            label = "P = 0.094", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = 29, xend = 1, yend = 30)) + 
  geom_segment(size = .8,
               aes(x = 5, y = 29, xend = 5, yend = 30))
```

@@

## Plotting body condition change with the interaction second infection intensity - infection stage
## -----------------------------------------------------------------------------
tolres <- 
  ggplot(data = filter(df, !Infection_intensity.2 %in% c(NA,0), !Diff_Inf_Class %in% NA),
         aes(y = delta_body_condition, x = as.numeric(Infection_intensity.2), col = Diff_Inf_Class)) + 
  geom_jitter(size = 3, alpha = 0.5) + 
  geom_smooth(alpha = 0.2, size = 1.5, method = 'lm',se = T, width = 0.2) + 
  facet_wrap(facets = 'Diff_Inf_Class') +
  labs(y = "Body condition change",x = "Parasitemia \n at second sampling",
       col = 'Infection stages') + 
  scale_x_continuous(breaks = seq(2,5,1),labels = c("1","2","3","4")) + 
  scale_color_manual(labels = c("Increasing", "Peak", "Decreasing"),
                     values = c("#3B9AB2", "#E1AF00", "#F21A00")) + 
  theme_pubr() +
  annotate(geom = 'segment', y = -Inf, yend = Inf, color = 'black', x = Inf, xend = Inf, size = 0.5) +
  annotate(geom = 'segment', y = -Inf, yend = Inf, color = 'black', x = -Inf, xend = -Inf, size = 0.5) +
  theme(legend.position = "none", 
        axis.text = element_text(size = 18), 
        axis.title = element_text(size = 20),
        strip.text = element_text(face = "bold", size = 16))

@@

ggsave(g,filename = "./Figs/Fig3_delta_body_condition_to_parasitemia_second.png",units = "px", width = 3700, height = 2800)
```

>>>>

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5),size=5) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) + 
  scale_x_discrete(labels = c("First\nsampling","Second\nsampling","First\nsampling","Second\nsampling"))+  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 26, face = "bold"),
        legend.text = element_text(size = 24),
        legend.title = element_blank(),
        legend.position = c(0.85, 0.90),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text.x = element_text(size = 24, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 24, margin =  unit(c(0, 1, 0, 0), "mm")))

@@

## Figure 2b
## -----------------------------------------------------------------------------
Fig_2B <- ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1)) +
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey")) + 
  scale_x_discrete(labels = c("First\nsampling","Second\nsampling"))+
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 24),
        legend.title = element_text(size = 26),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 5, 0, 0), "mm")),
        axis.text.x = element_text(size = 26, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 26, margin =  unit(c(0, 1, 0, 3), "mm")),
        plot.margin = unit(c(0,0,0,30), "pt"))

@@

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.3),
            label = "t = -2.80, df = 140, P = 0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.1)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.1,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

@@

Fig_4C = ggplot(mod_bc_para) + 
  aes(as.factor(mod_bc_para$x),mod_bc_para$predicted) + 
  geom_hline(yintercept = mean(mod_bc_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc_para$lower_CI, ymax = mod_bc_para$higher_CI),
                colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc_para$predicted - mod_bc_para$se,
                    ymax = mod_bc_para$predicted + mod_bc_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),  
        axis.text = element_text(size = 16), 
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 60, xend = 5, yend = 60)) + 
    geom_text(aes(x = 3, y = 67),
              label = "t = -1.76, df = 247, P = 0.080", size = 5) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 60, xend = 1, yend = 53)) + 
  geom_segment(size = .8,
                 aes(x = 5, y = 60, xend = 5, yend = 53))
```

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4D = ggplot(mod_gr_para) + 
  aes(as.factor(mod_gr_para$x),mod_gr_para$predicted) + 
  geom_hline(yintercept = mean(mod_gr_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr_para$lower_CI,
                    ymax = mod_gr_para$higher_CI), colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr_para$predicted - mod_gr_para$se,
                    ymax = mod_gr_para$predicted + mod_gr_para$se),size = 8, 
                colour = wes_palette("Zissou1",5), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
               aes(x = 1, y = 30, xend = 5, yend = 30)) + 
  geom_text(aes(x = 3, y = 31),
            label = "t = 2.06, df = 258, P = 0.094", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = 29, xend = 1, yend = 30)) + 
  geom_segment(size = .8,
               aes(x = 5, y = 29, xend = 5, yend = 30))
```

@@

## Plotting body condition change with the interaction second infection intensity - infection stage
## -----------------------------------------------------------------------------
tolres <- 
  ggplot(data = filter(df, !Infection_intensity.2 %in% c(NA,0), !Diff_Inf_Class %in% NA),
         aes(y = delta_body_condition, x = as.numeric(Infection_intensity.2), col = Diff_Inf_Class)) + 
  geom_jitter(size = 3, alpha = 0.5) + 
  geom_smooth(alpha = 0.2, size = 1.5, method = 'lm',se = T, width = 0.2) + 
  facet_wrap(facets = 'Diff_Inf_Class') +
  labs(y = "Body condition change",x = "Parasitemia \n at second sampling",
       col = 'Infection stages') + 
  scale_x_continuous(breaks = seq(2,5,1),labels = c("1","2","3","4")) + 
  scale_color_manual(labels = c("Increasing", "Peak", "Decreasing"),
                     values = c("#3B9AB2", "#E1AF00", "#F21A00")) + 
  theme_bw()+
  annotate(geom = 'segment', y = -Inf, yend = Inf, color = 'black', x = Inf, xend = Inf, size = 0.5) +
  annotate(geom = 'segment', y = -Inf, yend = Inf, color = 'black', x = -Inf, xend = -Inf, size = 0.5) +
  theme(legend.position = "none", 
        axis.text = element_text(size = 18), 
        axis.title = element_text(size = 20),
        strip.text = element_text(face = "bold", size = 16))

@@

ggsave(g,filename = "./Figs/Fig3_delta_body_condition_to_parasitemia_second.png",units = "px", width = 3700, height = 2400)
```
